Development of a educational and sustainable robotic platform for children and adults

### 2013-12-24 doc/paper.tex
autocommited by makefile

@@

@@

\subsubsection{Infrared}
Infrared based data transfer is widely used in remote control application where a small amount of data is transferred. Infrared communication is easily interfered by other infrared signals. This limits the data rate since error correction mechanisms must be implemented.

When using infrared signals only one robot can send at a time which requires collision detection. The hardware requirements for infrared transmissions are ver low, only a infrared reciever and a infrared LED is required which sums up to a total of 0.5 USD. The infrared reciever and the infrared LED require one free GPIO pin on the microcontroller each.

@@

Attaching the nRF24L01 to the microcontroller requires four SPI pins (MISO, MOSI, SCK and CS), an additional CE (chip enable) pin and an external interrupt pin to indicate state. Even though the microcontroller runs at 3.3V the inputs are 5V tolerant.

Two open source libraries Mirf and RF24 exist where RF24 is licensed under a GPL 2.0 license and Mirf is provided as is. 
\subsubsection{Conclusion}

\subsection{Microcontroller}
Depending on the microcontroller, the final robot has different processing capabilities. Its capabilities are mainly influenced by the processor frequency which influences the number of instructions per second, the interfaces such as I2C or SPI and the amount of flash storage and RAM.

@@

>>>>

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.5\textwidth]{images/30_ultrasonic.jpg}
  \caption{HC-SR04 ultrasonic distance sensor (CC-BY-SA Ben Oswald)}
\end{figure}

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.5\textwidth]{images/30_compass.jpg}
  \caption{Triple axis compass on breakout-board (CC-BY-SA Ben Oswald)}
\end{figure}

@@

\subsubsection{Infrared}
Infrared based data transfer is widely used in remote control application where a small amount of data, usually a single word, is transferred. Infrared communication is easily interfered by other infrared signals. This limits the data rate since error correction mechanisms must be implemented.

\begin{figure}[h!]
  \centering
  \begin{subfigure}{0.48\textwidth}
  \centering
  \includegraphics[width=0.9\linewidth]{images/30_irdiodes.jpg}
  \caption{Infrared diode (sender)}
  \end{subfigure}
  \begin{subfigure}{0.48\textwidth}
  \centering
  \includegraphics[width=0.9\linewidth]{images/30_ir.jpg}
  \caption{Infrared receiver (reciever)}
  \end{subfigure}
  \caption{Components for infrared communication (CC-BY-SA Ben Oswald)}
\end{figure}

When using infrared signals only one robot can send at a time which requires collision detection. The hardware requirements for infrared transmissions are ver low, only a infrared reciever and a infrared LED is required which sums up to a total of 0.5 USD. The infrared reciever and the infrared LED require one free GPIO pin on the microcontroller each. To ease programming the reciever should be attached to a microcontroller pin which is suitable for external interrupts.

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.5\textwidth]{images/30_nrf24l01.jpg}
  \caption{nRF24L01 breakout board (CC-BY-SA Ben Oswald)}
\end{figure}

Attaching the nRF24L01 to the microcontroller requires four SPI pins (MISO, MOSI, SCK and CS), an additional CE (chip enable) pin and an external interrupt pin to indicate state. Even though the microcontroller runs at 3.3V the inputs are 5V tolerant. The maximum data rate of the module is 2 Mbit/s, lower data rates (1 Mbit/s or 250Kbit/s) increase the maximum distance between two modules.

Two open source libraries Mirf and RF24 exist. RF24 is licensed under a GPL 2.0 license and Mirf is provided under a permissive "as is" license. The price for a nRF24L01 on a breadout-board is around 3 USD on ebay, other modules with external antenna have a higher price tag.
\subsubsection{Conclusion}
In terms of cost-effectiveness the nRF24L01 is the best solution for wireless communication between robots. The transmission is robust, easy to implement due to library support. Since the chip comes mounted on a breadboard with 2x4 header, a user can populate a socket and add the module on demand.
\subsection{Microcontroller}
Depending on the microcontroller, the final robot has different processing capabilities. Its capabilities are mainly influenced by the processor frequency which influences the number of instructions per second, the interfaces such as I2C or SPI and the amount of flash storage and RAM.

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.5\textwidth]{images/30_lithium-polymer.jpg}
  \caption{Lithium-polymer battery (CC-BY-SA Ben Oswald)}
\end{figure}
